r5r: rapid realistic routing on multimodal transport networks with R5 in R

### 2020-11-11 r-package/paper2020/paper2020.Rmd
changes to paper text

---
title: 'r5r: rapid realistic routing on multimodal transport networks with R5 in R'
author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
abstract: "Routing is a key step in transport planning. Nonetheless, researchers and practitioners still face challenges regarding this task, such as the cost of licensed software and the long computation times to run complex scenarios. Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single origin/destination pair. This paper describes r5r, an open source R package built on top of R<sup>5</sup> that allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense. The development of this package contributes to the transport field as it offers an easy-to-use powerful tool for robust planning and research."
urlcolor: blue
bibliography: references.json
---

@@

-   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>. This function (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.

@@

The package can be installed with the following command:

@@

The first step when using r5r is to set the amount of memory available to Java. By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which is not enough for the tasks r5r is designed to carry out, such as calculating travel time matrices on large multimodal networks. To increase available memory to 2GB[^1], for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:

[^1]: Note that large study areas may require significantly more memory to run properly.

>>>>

---
title: 'r5r: rapid realistic routing on multimodal transport networks with R5 in R'
author: "Rafael H. M. Pereira, Marcus Saraiva, Daniel Herszenhut, Carlos Kaue Braga"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
abstract: "Routing is a key step in transport planning. Nonetheless, researchers and practitioners still face challenges to perform this task due the costs of licensed software or the long computation times to run multiple routing analysis in complex transport networks. Conveyal's R<sup>5</sup> is a multimodal transport network router that offers meaningful planning features, such as accounting for service reliability issues and returning multiple travel time estimates and itineraries for a single or multiple origin/destination pairs. This paper describes r5r, an open source R package that leverages on R<sup>5</sup> to allow users to efficiently compute travel time matrices and generate detailed itineraries between a set of origins and destinations at no expense. The development of this package contributes to the transport field as by offerin an easy-to-use powerful tool for robust planning and research."
urlcolor: blue
bibliography: references.json
---

@@

-   `setup_r5()`: builds a multimodal transport network used for routing in R<sup>5</sup>. This function automatically (1) downloads/updates a compiled R<sup>5</sup> JAR file, tailored for the purposes of the package, and stores it locally in the r5r package directory for future use; and (2) combines the OSM and GTFS data sets to build a routable network object.

@@

The package can be installed with the following command[^1]:

@@

The first step when using r5r is to set the amount of memory available to Java. By default `R` allocates only 512 MB of memory to the Java Virtual Machine (JVM), which might not be enough to perform tasks on large multimodal networks. To increase available memory to 2GB, for example, the user needs to set the `java.parameters` option at the beginning of the script as follows:

[^1]: To use r5r, one needs to have locally installed Java SE Development Kit, version 11.0.8 or higher. Available at <https://www.oracle.com/java/technologies/javase-jdk11-downloads.html>.
